What Can Stack Traces Tell us about Exceptions in Java?

### 2014-07-07 android-stacks.tex
improving process description

Hence, in such context where we have access to wealth of information available
on open-source repositories such as GitHub, some questions arise: What are the common
characteristics of Github crash reports? How can we benefit from such cross-project crash analysis?

@@

%\subsubsection{Exception Throwing and Handling}

In Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language (e.g., out-of-bounds array index,
division-by-zero, access to a null reference); or (iii) implicitly thrown by
the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

%When an exception is signaled it causes a transfer of control from the point
%where the exception occurred to a point that can be specified by the programmer
%(exception handler) - in Java it is represented by the try-catch block .
%The search for the handler to deal with a raised
%exception occurs along the dynamic invocation chain. 
%If no handler is found in the chain, the default handler associated to the 
%Thread or ThreadGroup is invoked, which usually terminates the system 
%and prints the exception stack trace.

%\subsubsection{Exception Chaining} Exception Wrapping 
%This stack trace was extracted from an issue of an Android application in
%Github.

The Exception wrapping (also known as exception chaining) is a common way of 
propagating exceptions in Java programs, 
which allows one exception to be wrapped in another exception and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

This work describes an exploratory study which was based on a sequential
mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
integrating both quantitative and qualitative evidence along the stages of the
research process. The main goal of this study was to gain a better understanding
of the use of exceptions in Java programs based mainly on the information
available on exception stack traces. More specifically, we aimed at answering
the following research questions:

@@

  \item[RQ1] What is the prevalence of stack traces on  reported issues?
    %\item[RQ2] What are the most common exceptions?

  \item[RQ2] What are the common characteristics of such stack traces?

  \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace
    analysis?

@@

For RQ1 and RQ2, we explore the domain quantitatively, and use their answers as
a basis to investigate the third research question. RQ3 aims at examining whether
exception (mis)-uses could be revealed through stack trace analysis. To
support this investigation, source code and bytecode analysis was used to
leverage the  understanding of stack traces. 
%For RQ3 we investigate
%quantitatively and highlight interesting cases by exploring cases qualitatively.
Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
conducted to answer these questions.

@@

\subsection{Github Data}
\label{sec:git}

@@

>>>>

Hence, in such context where we have access to wealth of information available
on open-source repositories such as GitHub, some questions arise: What are the common
characteristics of Github crash reports?

@@

Hence, in Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language (e.g., out-of-bounds array index,
division-by-zero, access to a null reference); or (iii) implicitly thrown by
the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

When an exception is throw it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block .

A common way of  propagating exceptions in Java programs, is the exception wrapping (also known as exception chaining), which allows one exception to be wrapped in another exception and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

This work describes an exploratory study which was based on a sequential
mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
integrating both quantitative and qualitative evidence along the stages of the
research process. The main goal of this study was to gain a better understanding
of the use of exceptions in Java programs based mainly on the information
available on exception stack traces. The exceptions available on exception stack trace
are not limited to the exceptions expliciltly thrown by the Java developer but also include
the exceptions implicitly thrown by the JVM during the program execution,
due to a programming error (e.g., out-of-bounds array index, division-by-zero, 
access to a null reference) or an internal error or resource limitation (e.g.,
OutofMemoryError).

 More specifically, we aimed at answering
the following research questions:

@@

  \item[RQ1] What are the common characteristics of such stack traces?

  \item[RQ2] What patterns of exception (mis)-use emerge from the stack trace
    analysis?

@@

For RQ1 and RQ2, we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively. To support this investigation, source code and bytecode 
analysis were used in combination 
with manual inspection to to everage the  understanding of stack traces.

%For RQ3 we investigate
%quantitatively and highlight interesting cases by exploring cases qualitatively.
Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
conducted to answer these questions.

@@

\subsection{Study Context and Github Data}
\label{sec:git}

 What is the prevalence of stack traces on reported issues?
   What are the most common exceptions?

@@

TO DISCUSS: How can we benefit from such cross-project crash analysis?
